---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MBGapp

<!-- badges: start -->
<!-- badges: end -->

**MBGapp** is an interactive Shiny application for teaching and practising
model-based geostatistics (MBG). It guides users through the complete spatial
analysis workflow — data exploration, variogram fitting, model estimation, and
spatial prediction — without requiring any coding.

## Features

- **Three outcome types** — continuous, prevalence (binomial), and count (Poisson)
- **Interactive maps** — leaflet-based exploration and prediction maps with pan/zoom
- **Empirical variogram** — adjustable bins, distance cutoff, and correlation functions
- **Bayesian estimation** — geostatistical model fitting via RiskMap (MCMC) with an
  optional INLA (fast Bayes) backend when the INLA package is installed
- **Spatial prediction** — mean surface, standard error, exceedance probability, and
  quantile maps over a user-defined grid
- **Downloadable report** — generate a PDF report with selected outputs
- **Shapefile support** — upload a boundary shapefile to constrain maps and grids
- **Covariate support** — include linear and non-linear covariate effects

## Workflow

| Tab | What you do |
|-----|-------------|
| **Explore** | Upload data, choose data type, inspect the spatial distribution on an interactive map and scatter plots |
| **Variogram** | Examine spatial correlation structure; fit theoretical variogram models |
| **Estimation** | Fit a geostatistical model; view parameter estimates and 95% confidence intervals |
| **Prediction** | Map the predicted surface over the study region |
| **Report** | Download a PDF report of selected outputs |

## Installation

Install the development version from GitHub:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("olatunjijohnson/MBGapp", ref = "main")
```

Then launch the app:

```{r eval=FALSE}
library(MBGapp)
run_app()
```

## Run without installing

```{r eval=FALSE}
shiny::runGitHub(
  repo     = "MBGapp",
  username = "olatunjijohnson",
  ref      = "main",
  subdir   = "inst/MBGapp"
)
```

## Online version

Access the app directly in your browser — no R installation needed:

**<https://olatunjijohnson.shinyapps.io/mbgapp/>**

## Example data

The package ships with the **Loa loa** prevalence survey dataset from Cameroon
(columns: `Longitude`, `Latitude`, `Positive`, `Examined`). A 10 km prediction
grid and covariate rasters for Cameroon are also included.

Additional example files can be downloaded from Google Drive:

- [Loa loa survey data](https://drive.google.com/uc?export=download&id=1nGRuw-UUFYbG0Wl4XD4noxaNb1iyyky9)
- [10 km prediction grid](https://drive.google.com/uc?export=download&id=1-P880pttLqOlyQnW6_K1MPsEQo_Gh4zM)
- [Covariate rasters](https://drive.google.com/uc?export=download&id=1fe5i08xIghPP_kHUHiL0SNT1usa8L-OD)

## Dependencies

MBGapp uses the following R packages:

`shiny`, `shinyjs`, `sf`, `terra`, `leaflet`, `leafem`, `tidyterra`, `stars`,
`ggplot2`, `dplyr`, `readr`, `tidyr`, `magrittr`, `splines`, `geoR`,
[RiskMap](https://github.com/claudiofronterre/RiskMap) (MCMC backend),
and optionally [INLA](https://www.r-inla.org) (fast Bayes backend).

## Authors

Olatunji Johnson, Claudio Fronterre, Emanuele Giorgi
CHICAS, Lancaster Medical School, Lancaster University
