---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MBGapp

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/olatunjijohnson/MBGapp.svg?branch=main)](https://travis-ci.org/olatunjijohnson/MBGapp)
[![Codecov test coverage](https://codecov.io/gh/olatunjijohnson/MBGapp/branch/master/graph/badge.svg)](https://codecov.io/gh/olatunjijohnson/MBGapp?branch=main)
<!-- badges: end -->

The goal of MBGapp is to allow user to explore and analyse geostatistical data

## Installation

<!-- You can install the released version of MBGapp from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("MBGapp") -->
<!-- ``` -->

You can install the development version on github with 

``` r
devtools::install_github("olatunjijohnson/MBGapp", ref="main")
```

## Example

Download an example dataset from [here](https://drive.google.com/uc?export=download&id=1nGRuw-UUFYbG0Wl4XD4noxaNb1iyyky9). The data consist of the Loa loa prevalence survey data in Cameroon. You can also download an example 10km by 10km grid locations in Cameroom and the corresponding covariates at the pixel-level from [here](https://drive.google.com/uc?export=download&id=1-P880pttLqOlyQnW6_K1MPsEQo_Gh4zM) and [here](https://drive.google.com/uc?export=download&id=1fe5i08xIghPP_kHUHiL0SNT1usa8L-OD), respectively.

This is a basic example which shows you how to run the app:

```{r eval=FALSE}
library(MBGapp)
## run the App
run_app()  # use the code
```

## Alternative way to run in R

You can also run the following line of code to run in R

```{r eval=FALSE}
shiny::runGitHub(repo="MBGapp", username= "olatunjijohnson", ref="main", subdir = "inst/MBGapp")
```


## Online version

The app can also be accessed online via the following link:

https://olatunjijohnson.shinyapps.io/mbgapp/
